---
title: "9 November 2022"
author: "Erika"
date: "11/9/2022"
output: html_document
---

In class today we are going to take a look at some steps that might make some of your Cap DAP work easier:

Using the tidyverse package and, in particular, noticing:
tidyr - moving between wide and long formats
stringr - finding and breaking up text strings
Dealing with dates
lubridate package
Saving data as .csv from R

Let's get started!

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Pull in some data
```{r}
temp <- read.csv(here("Data", "temp_by_year.csv"))
```

Some issues in the data:
1. Wide rather than long
2. Year with an X
3. Two variables in single plot column

We need tidyr and stringr to fix

Step 1 - wide to long format with tidyr

```{r}
temp_long <- pivot_longer(temp, cols = X1979:X1990,   names_to = "Year", values_to = "temp")
```
That worked, but introduced a new problem, which is that each year starts with X and thus can't be converted to a date. So, how do we get rid of the X?

One way that we have already learned is with `which`

For example:
```{r}
Years<-unique(temp_long$Year)
fix<-which(temp_long$Year == "X1979")
```
But let's look at a different way with stringr

There are multiple ways you could do this:
`str_replace`
`str_remove`
`str_split`

```{r}
temp_long1 <- str_replace(temp_long$Year, "X", "")

temp_long$newYear <- temp_long1
```
Now save the data frame

```{r}
write.csv(temp_long, here("Data", "Processed", "correct_temp_by_yearX.csv"))

write.csv(temp_long, here("Data", "Processed", "correct_temp_by_year.csv"), row.names = FALSE)
```

Let's make yet another year column to play with
```{r}
temp_long$Year3<-temp_long$newYear
```

Now make newYear into integer
```{r}
temp_long$newYear <- as.integer(temp_long$newYear)
```
Now let's convert Year3 into a date.

Use package lubridate

```{r}
library(lubridate)
```
Let's arbitrailty make year into 1/1/year

use `paste0` function

```{r}
temp_long$Year3 <- paste0("01-01-",temp_long$Year3)
```
Now let's try to make it a date

```{r}
temp_long$Year3 <- mdy(temp_long$Year3)
```












