---
title: "16Nov2022-ANOVA"
author: "Erika"
date: "2022-11-16"
output: html_document
---

Let's do a two-way ANOVA

```{r}
rm(list =ls())
library(tidyverse)
library(ggfortify)
library(here)
```

And now pull in some data about growth in cows

```{r}
DF <- read.csv(here("Data", "growth.csv"))
```
Cows were fed one of three diets (barley, oats, wheat) and the diets were enhanced with one of 4 supplements (control and a bunch of names). Each combination of diet and supplement were replicated.


Look at the data
```{r}
glimpse(DF)
```

So two of our variables are categorical: supplement and diet

Let's convert them to factor and check their levels
```{r}
DF$supplement <- as.factor(DF$supplement)
DF$diet <- as.factor(DF$diet)

levels(DF$supplement)
levels(DF$diet)
```

We'd like the control to be the reference factor, so let's relevel
```{r}
DF<- mutate(DF, supplement = relevel(supplement, ref = "control"))
levels(DF$supplement)
```

We need to look at the mean and growth rate of each of the combinations, so use group by and summarise

```{r}
sumDF <- DF %>% group_by(diet, supplement) %>% summarise(meanGrow = mean(gain))
```

Now let's make an interaction plot

Think about this statement: *the effect of supplement type on cow weight gain depends on the diet*

```{r}
ggplot(sumDF, aes(supplement, meanGrow))+
     geom_point()+
     theme_bw()
```
Now let's connect the dots!

Modify our ggplot thusly:
```{r}
ggplot(sumDF, aes(supplement, meanGrow, color = diet, group = diet))+
     geom_point()+
     geom_line()+
     theme_bw()
```
Now create the model

```{r}
model_cow <-lm(gain ~ diet*supplement, data = DF)
```
And check assumptions
```{r}
autoplot(model_cow, smooth.color = NA)
```
Now look at the model

```{r}
anova(model_cow)
```


And look at the summary

```{r}
summary(model_cow)
```

And now make our final plot

First get some different values for the plot
```{r}
sumDF<-DF %>% group_by(diet, supplement) %>% summarise(
     meanGrow = mean(gain),
     seGrow = sd(gain)/sqrt(n())#watch out for NAs and see next code block for a better version
)
```

```{r}
sumDF<-DF %>% group_by(diet, supplement) %>% summarise(
     meanGrow = mean(gain),
     seGrow = sd(gain)/sqrt(sum(!is.na(gain)))#There! Much better version
)
```

Now we can make our figure better:
```{r}
ggplot(sumDF, aes(supplement, meanGrow, color = diet, group = diet))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meanGrow - seGrow,
                       ymax = meanGrow + seGrow), width = 0.1)+
     theme_bw()
```

