---
title: "R Notebook"
output: html_notebook
---

Set things up

```{r}
rm(list = ls()) #clear R's brain
library(dplyr)
library(ggplot2)
library(here)
```

Now suck in some data

```{r}
DF <- read.csv(here("Data", "data_types_exercise.csv"))
str(DF)
```

And now with stringsAsFactors = TRUE to see what that does.

```{r}
DF2 <-read.csv(here("Data", "data_types_exercise.csv"),stringsAsFactors = TRUE)

str(DF2)
```

Let's create some non-data-frame data

```{r}
dogs<-c("Gus", "Tulip", "Emma", "Tex")
```

Make another variable called temps
```{r}
temps <- c(-15.2, 0.0, 3.8, -18.0, 1.5)
```
Can I use the str() function outside of a data frame?

```{r}
str(temps)
str(DF)
```
Now let's make a single variable data frame
```{r}
cats <- as.data.frame(c("Dewey", "Silver", "Bandon"))
str(cats)
```

How do we access certain columns of a dataframe?
Remember that $? Let's use it!

```{r}
#display the dogs variable
dogs
#display the Majors variable in the DF dataframe
DF$Majors
```
How do I access specific values from a data frame?

Let's make a small easy to use dataframe using cbind

```{r}
Farenheit <- c(70.3, 86.7, 42.5)
Centegrade <- c(20.6, 34.2, 17.1)
Places <- c("Canton", "Potsdam", "Madrid")
```
Now put them together into a new dataframe using cbind
```{r}
MyData <-as.data.frame(cbind(Centegrade, Farenheit, Places))
```

Now use indexing to find values

```{r}
MyData[2, 1]
MyData[1,3]
MyData[2,]
MyData[,3]
MyData$Places
MyData[2,1:2]
```





