---
title: "In Class 31 Oct 2022"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

We are working on exploratory data analysis using the data analysis workflow demonstrated in class.

Today in class students are using the soil_subset_outliers data set. Their first task is to:

1. Identify the types of variables in the dataset

2. Describe and visualize central tendency and dispersion

3. Identify any outliers and decide how to deal with them.

Let's get started:
```{r}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(here)

DF<-read.csv(here("Data", "soil_subset_outliers.csv"))
```
Let's begin by getting some summary statistics
```{r}
summary(DF)
```
Immediately I see that some variables are coming in as character that I'd like as factor. Let's fix that.

```{r}
DF$Forest<-as.factor(DF$Forest)
DF$Plot<-as.factor(DF$Plot)
DF$Sample<-as.factor(DF$Sample)
```
Now let's get the boxplots and histograms for the data

```{r}
DF %>% ggplot(aes(x = Organic.Matter..))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Nitrate.N.ppm))+
     geom_histogram()
```
Next
```{r}
DF %>% ggplot(aes(x = Base.Saturation.Ca))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Base.Saturation.Mg))+
     geom_histogram()
```

```{r}
DF %>% ggplot(aes(x = Base.Saturation.K))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Base.Saturation.Na))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Base.Saturation.H))+
     geom_histogram()
```

```{r}
DF %>% ggplot(aes(x = Base.Saturation.Total))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Exch.Acidity))+
     geom_histogram()
```
```{r}
DF %>% ggplot(aes(x = Buffer.pH))+
     geom_histogram()
```


