---
title: "Silly Stuff with R"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(ggplot2)
library(plotly)
library(gtools)
```

## Happy Solstice!

As you may know, December 21 is the shortest day of the year in the northern hemisphere and is called the [winter solstice](https://en.wikipedia.org/wiki/Winter_solstice).

Let's celebrate the solstice with some R code!


Follow through each of these sections to learn something new. 
### 1. The diverging bar plot

Create a data frame 
```{r}
#   - specify 2*n bars (but have only n unique values)
#   - set divergence values (values at the same level should differ only in the sign)
#   - set labels for different parts of the plot

df <- data.frame("wish" = c("YS",  "YS", "IDA", "IDA", "HOL", "HOL",
                            "PY", "PY", "HAP", "HAP"),
                 "pos" = c(0.75, -0.75, 3.5, -3.5, 2.5, -2.5,
                           1.5, -1.5, 0.3, -0.3),
                 "part" = c(rep("bottom", 2), rep("tree", 6), rep("star", 2)))
```

Now get ready to plot
```{r}
# Convert wish to factor, specify levels to have the right order
df$wish <- factor(df$wish, levels = c("YS", "IDA", "HOL", "PY", "HAP"))

ggplot(df, aes(x = wish, y = pos, fill = part)) + geom_bar(stat="identity") +
    coord_flip() +
    theme_minimal() +
    ylim(-5, 5) +
    theme(legend.position = "none",
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    scale_fill_manual(values = c("#643413", "#FDBA1C", "#1A8017")) +
    geom_point(aes(x=3.7, y=0.5), colour="#CF140D", size=12) +
    geom_point(aes(x=2.5, y=-1), colour="#393762", size=12) +
    geom_point(aes(x=1.7, y=1.5), colour="#CF140D", size=12) +
    geom_point(aes(x=2.8, y=2.5), colour="#393762", size=12) +
    geom_point(aes(x=1.8, y=-2.8), colour="#CF140D", size=12)
```
2. The Dirichlet distrubtion

```{r}
library(gtools)

set.seed(24) # seed, so our tree is reproducible!

n_tree <- 7000
n_ornaments <- 50

# Sample for tree and ornaments
#   - sample n_tree/n_ornaments points from the tree spiral
#   - alpha = c(1, 1, 1) -> points are uniformly distributed
#   - map values from R^3 to R^2
dir_samples_tree <- rdirichlet(n_tree, c(1, 1, 1))
df <- data.frame("x" = dir_samples_tree[, 2] - dir_samples_tree[, 3],
                 "y" = dir_samples_tree[, 1],
                 "size" = 1,
                 "type" = "tree")

dir_samples_ornaments <- rdirichlet(n_ornaments, c(1, 1, 1))
df <- rbind(df, data.frame("x" = dir_samples_ornaments[, 2] - dir_samples_ornaments[, 3],
                           "y" = dir_samples_ornaments[, 1],
                           "size" = 2,
                           "type" = "ornaments"))

ggplot(df, aes(x=x, y=y, color = type)) +
    geom_point(aes(size=size)) +
    scale_color_manual(values = c("#CF140D", "#1A8017")) + # set colors
    theme_void() + # remove axes, labels, ticks etc.
    theme(legend.position = "none") + # remove legend
    scale_size_continuous(range = c(1.5, 8)) + # set scale for points
    ylim(0, 1.1) + xlim(-1.2, 1.2) +
    annotate("text", x = 0, y = 1.1, label = "Happy Solstice!",
             color = "#CF140D", size = 7,
             family = "Palatino", fontface = "bold.italic") +
    geom_text(aes(x=0, y=1), label="★", size=15, colour="#FDBA1C",
              family = "HiraKakuPro-W3") # font that supports the character
```


```{r}
library(plotly)

set.seed(24)

n_tree <- 1000
n_ornaments <- 20
n_lights <- 300

# Generate spiral data points
x <- c()
y <- c()
z <- c()

for (i in 1:n_tree) {
    r <- i / 30
    x <- c(x, r * cos(i / 30))
    y <- c(y, r * sin(i / 30))
    z <- c(z, n_tree - i)
}

tree <- data.frame(x, y, z)

# Sample for ornaments:
#   - sample n_ornaments points from the tree spiral
#   - modify z so that the ornaments are below the line
#   - color column: optional, add if you want to add color range to ornaments
ornaments <- tree[sample(nrow(tree), n_ornaments), ]
ornaments$z <- ornaments$z - 50
ornaments$color <- 1:nrow(ornaments)

# Sample for lights:
#   - sample n_lights points from the tree spiral
#   - Add normal noise to z so the lights spread out
lights <- tree[sample(nrow(tree), n_lights), ]
lights$x <- lights$x + rnorm(n_lights, 0, 20)
lights$y <- lights$y + rnorm(n_lights, 0, 20)
lights$z <- lights$z + rnorm(n_lights, 0, 20)

# hide axes
ax <- list(
    title = "",
    zeroline = FALSE,
    showline = FALSE,
    showticklabels = FALSE,
    showgrid = FALSE
)

plot_ly() %>%
    add_trace(data = tree, x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "#1A8017", width = 7)) %>%
    add_markers(data = ornaments, x = ~x, y = ~y, z = ~z,
                type = "scatter3d",
                marker = list(color = ~color,
                              colorscale = list(c(0,'#EA4630'), c(1,'#CF140D')),
                              size = 15)) %>%
    add_markers(data = lights, x = ~x, y = ~y, z = ~z,
                type = "scatter3d",
                marker = list(color = "#FDBA1C", size = 3, opacity = 0.8)) %>%
    layout(scene = list(xaxis=ax, yaxis=ax, zaxis=ax), showlegend = FALSE)
```
The idea for this project has been modified from [this web page](https://towardsdatascience.com/christmas-cards-81e7e1cce21c).
