---
title: "9 November 2022"
author: "Erika"
date: "11/9/2022"
output: html_document
---

In class today we are going to take a look at some steps that might make some of your Cap DAP work easier:

Using the tidyverse package and, in particular, noticing:
tidyr - moving between wide and long formats
stringr - finding and breaking up text strings
Dealing with dates
lubridate package
Saving data as .csv from R

Let's get started!

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Pull in some data
```{r}
temp <- read.csv(here("Data", "temp_by_year.csv"))
```

Some issues in the data:
1. Wide rather than long
2. Year with an X
3. Two variables in single plot column

We need tidyr and stringr to fix

Step 1 - wide to long format with tidyr

```{r}
temp_long <- pivot_longer(temp, cols = X1979:X1990,   names_to = "Year", values_to = "temp")
```
That worked, but introduced a new problem, which is that each year starts with X and thus can't be converted to a date. So, how do we get rid of the X?

One way that we have already learned is with `which`

For example:
```{r}
Years<-unique(temp_long$Year)
fix<-which(temp_long$Year == "X1979")
```
But let's look at a different way with stringr

There are multiple ways you could do this:
`str_replace`
`str_remove`
`str_split`









