---
title: "Practicing with dplyr"
output: html_notebook
---

Set things up
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
```

Pull in the data

```{r}
DF <-read.csv(here("Data", "tomato_data.csv"), stringsAsFactors = TRUE)
```

Grab just the number of flowers column
```{r}
select(DF, Num_flowers)
```
Get 3rd through 8th rows of data
```{r}
slice(DF, 3:8)
```
Get just the control data
```{r}
filter(DF, Treatment == "Control")
```
Use piping to do the following:

Create a new variable called "big_toms" that returns a data frame consisting of a single column of only the plant numbers from plants that produced fruit with a mean weight of 150 g or more
```{r}
big_toms<-DF %>% filter(Avg_tomato_weight_g >=150) %>% select(Plant_number)
```

Use group_by and summarize to determine the mean number of flower produced in the control and treatment groups
```{r}
DF %>% group_by(Treatment) %>% summarize(meanFlowers = mean(Num_flowers))
```
Use group_by and summarize to determine the mean and standard deviation in mean fruit size (grams) between control and fertilized plants

```{r}
TomatoSizes <- DF %>% group_by(Treatment) %>% summarize(
  meanSizeGrams = mean(Avg_tomato_weight_g),
  sdSizeGrams = sd(Avg_tomato_weight_g)
)
```


